Modeling defoliation as a proxy for tree health: A case study using machine-learning and hyperspectral remote sensing data

### 2019-06-16 code/98-paper/journal/manuscript.tex
rewrite introduction

% Explain how remote sensing is used in forestry (potential to map forest health)
% Link remote sensing and machine-learning
The use of \ac{ML} algorithms for analyzing remote sensing data has seen a huge increase in the last decade \citep{lary2016}.
Especially since the launch of the first Sentinel satellite in 2014 the amount of freely available imagery data has increased greatly.
At the same time, the implementation and usability of algorithms (both statistical and machine-learning) has been greatly simplified in the open-source community.
Scientists can nowadays relatively easily feed large amounts of information into complex models in a semi-automated way.
These points enable the possibility to analyze (temporal) changes of the environment on a large scale \citep{ma2015}.

% link to forest health analysis and show exemplary studies
The monitoring of forest ecosystems is one exemplary field that profits from these new possibilities.
Specifically the monitoring of forest health profited from the ability of analyzing large scale data with efficient algorithms able to predict to unknown areas \citep{hawrylo2018, mascaro2014}.
For this task, usually optical data from multi-/hyperspectral satellites is used.
Various indices are calculated from the data with the aim of carrying out temporal change detections \citep{zhang2016} or conduct analyses about the current health status \citep{townsend2012}.
Vegetation indices have shown the potential to provide valuable information for analyzing triggers to forest health \citep{jiang2014, adamczyk2015}.
Most studies focus on applying the Random Forest algorithm \ac{RF} \citep{belgiu2016, lary2016, michez2016}.
Others are only used in few studies such as Support Vector Machines \citep{clark2018} or deep-learning algorithms such as Artifical Neural Networks \citep{ingram2005, rocha2018}.
Overall, the idea is to fit a robust model which can be used to make predictions to (large) unknown areas, providing valuable information about the health condition of forest stands without the need for ground-truth data in these regions.

% now mention why this study is important
In forest healh studies, it is valuable to have a lot input variables.
When searching for possible triggers to forest health within forests, it is unclear upfront which information from the data might reveal important relationships during model fitting.
If hyperspectral data is available, lots of indices can be derived.
One issue that comes with this approach is the case of "high-dimensionality" \citep{trunk1979, xu2016}.
Even though \ac{ML} algorithms are capable of handling highly-correlated input variables, the fitting time of models increases substantially and the interpretation of results becomes more complicated.
Also, in recent years the awareness of conducting proper tuning of model hyperparameters has increased in applied modelling fields.
The reseach questions of this study are formulated around the handling of feature-rich datasets in environmental modeling:

\begin{itemize}
	\item Do different environmental feature sets show a substantial different in performance when predicting defoliation?
    \item How do feature-selection methods perform amongst each other and do they improve the predictive performance?
    \item Is is possible to map defolation at a tree level using high-resolution hyperspectral data at a sufficient error rate?
    \item Which features were most important to the model to detect defoliation at trees?\end{itemize}

\noindent State-of-the-art machine-learning techniques were compared using three feature sets with different feature-selection approaches including a relatively novel approach of ensemble filters.
To model the numeric response variable defoliation at trees model-based optimization was applied for hyperparameter tuning.
Spatial block \ac{CV} was used to account for spatial autocorrelation in the data.

>>>>

% Explain how remote sensing is used in forestry (potential to map forest health)
% Link remote sensing and machine-learning
\noindent The use of \ac{ML} algorithms for analyzing remote sensing data has seen a huge increase in the last decade \citep{lary2016}.
Especially since the launch of the first Sentinel satellite in 2014 the amount of freely available imagery data has increased greatly.
At the same time, the implementation and usability of algorithms (both statistical and machine-learning) has been greatly simplified in the open-source community.
Scientists can nowadays relatively easily feed large amounts of information into complex models in a semi-automated way.
These points enable the possibility to analyze (temporal) changes of the environment on a large scale \citep{ma2015}.

% link to forest health analysis and show exemplary studies
Machine-learning methods in combination with remote-sensing data are used in many environmental fields such as vegetation cover analysis or forest carbon storage mapping \citep{mascaro2014, urban2018}.
The advantages of conducting analysis at large scale make these points a promising toolset also for forest health analysis.
This research field aims to enhance the understanding of biotic and abiotic triggers, for example by analyzing defoliation at trees \citep{hawrylo2018}.
Different approaches exist including temporal change detections \citep{zhang2016} or analysis about the current health status of forests on a stand level \citep{townsend2012}.
Defoliation at trees is an information which is able to serve as a proxy for forest health by describing the impact of biotic and abiotic pest triggers \citep{townsend2012, goodbody2018}.
Vegetation indices have shown the potential to provide valuable information when analyzing such scenarios \citep{jiang2014, adamczyk2015}.
Since both triggers and tree types are highly variable across regions, it cannot be known apriori which indices are most important for describing the actual relationship.
This emphasizes the need to extract as much information as possible from the available input data to generate features that can possibly explain the modeled relationship.
Besides \ac{VI} a less known type of indices that can be calculated from remote-sensing products are \ac{NRI}.
\ac{NRI} are arbitrary combinations of band information from remote-sensing data.
Especially when working with hyperspectral data, \ac{NRI} provide the possibility to calculate hundreds of indices that can be used during modeling.

% now that we introduced the importance of remote sensing indices for forest health,

Despite its popularity in environmental modeling, there are no studies so far that used machine-learning algorithms in combination with remote-sensing data to analyze defoliation on a tree level.
This study aims at closing this gap by analyzing defoliation at trees in northern Spain using airborne hyperspectral data.
To cope with the state-of-the-art methodology in modeling, a methodology which combines feature-selection and hyperparameter tuning across multiple algorithms is used.
The approach of generating many features comes along with the introduction of "high-dimensionality" into the analysis \citep{trunk1979, xu2016}.
Even though \ac{ML} algorithms are capable of handling highly-correlated input variables, the fitting time of models increase substantially and the interpretation of results becomes more complicated.
Besides modeling defoliation at trees this work also aims at providing an exemplary analysis of how to deal with spatial feature-rich data derived from a hyperspectral remote-sensing source.

The reseach questions of this study are the following:

\begin{itemize}
	\item Do different environmental feature sets show substantial differences in performance when predicting defoliation at trees?
    \item How are different feature-selection methods performing and do they improve the predictive performance of the final model?
    \item Is it possible to map defoliation at a tree level using high-resolution hyperspectral data at a sufficient error rate?
    \item Which features were most important to the model to detect defoliation at trees?\end{itemize}

\noindent State-of-the-art machine-learning techniques were compared using three feature sets with different feature-selection approaches including a relatively novel approach of ensemble filters.
Model-based optimization was applied for hyperparameter tuning of the chosen algorithms.
Spatial block \ac{CV} was used to account for spatial autocorrelation in the data.
